---
title: "Churn and Default"
author: "Jacob Ford, Solstice Data Scientist"
date: "`r Sys.Date()`"
output: 
  rmdformats::readthedown:
    toc_depth: 5

knit: (function(inputFile, encoding) { 
        rmarkdown::render(inputFile, encoding = encoding, 
        output_file = file.path(dirname(inputFile), 'churn_and_default.html')) })
---












# Introduction

**to-do: add intro**

## Literature Review

**to-do: to be added/folded into the Introduction to frame the macro-scope of this study**



**note**

*Analysis is awaiting Experian appending.* 


This analysis summaries the Nautilus customer data from the Frog Hollow and Howell community solar farm participation from January 2020 to April 2022. The goal of the analysis is to first describe the characteristics of the residents, the prevalence of default and/or churn rates, the prevalence of churn rates, and any statistically significant differences between groups in their default, late payments or churn rates.  


# Methods

The data is comprised of monthly payment performance and, where available, demographic data for residents. 32,385 monthly observations over 813 homes, 812 accounts and 621 users were initially observed. To avoid data quality issues and to analyze any potential churn and default trends, for Frog Hollow households the analysis will drop values in October 2021 until April 2022. This narrowed dataset is comprised of 31,704 observations; the homes, accounts and users did not change.  




```{r message=FALSE, warning=FALSE, include=FALSE}

#Load packages we'll need throughout the entire document

library(readxl)
library(sf)
library(dplyr)
library(janitor)
library(dplyr)
library(tidyverse)
library(nngeo)
library(plotly)
library(kableExtra)
library(haven)
library(rms)
library(rmdformats)
library(gtsummary)
library(ggplot2)
library(plotly)


options(scipen = 100)


```

```{r message=FALSE, warning=FALSE, include=FALSE}

## Change the below to match your local drive ##

input_folder <- '/Volumes/GoogleDrive/Shared drives/*Solstice | Low Income Inclusion/Projects/2019-2022 DOE SETO Contract Innovations/DOE SETO Project Implementation/Project Management/SETO Data Analysis/'


setwd(input_folder)


```



```{r message=FALSE, warning=FALSE, include=FALSE}

# 'raw' contains the monthly enrollment records for the sample solar farm - contains various account specific information

# There are three main ID variables to consider: user (user_id),  household (property_id) and utility account (utility_acct_number); we decided to group the analysis by accounts. The reason is simply that what we are measuring is rate of churn/default, and the account is the unit that performs these actions; not the user which may hold multiple accounts. Household ID and utility account are very similar, but for consistency, we hearafter focus on utlity account data



raw <- read_excel(paste(input_folder, "3. Churn and Default/Central Hudson Preliminary Analysis/Nautilus Raw Data.xlsx", sep=""), 
    sheet = "platform billing data")


#  'default_tag' - trigger that captures reason why an individual leaves a solar farm ; for non-churn/non-default records, it should be set to 0 (zero), not 'none' just for consitency


raw$default_tag[raw$default_tag=="none"] <- 0


raw <- raw %>%
  #filter particular case that was entered in error
  filter(user_id!=8377) %>%
  # make tag for frog hollow properties after 9/2021; assuming the entries for april 2022 are legit, this drops October 2021 values 
  mutate(FH_bill_issuetag = case_when(
         solar_farm_name=="Frog Hollow (CHGE)" & year == 2021 & month > 9 ~ 1,
         TRUE ~ 0),
         
         # Create dummy variable for default, the more extreme case of churn; every default is a churn, not ever churn is a default of course
         default_dummy = case_when(
           default_tag !=0 & default_tag !="Defaulted Payment" ~ 1,
           TRUE ~ 0),
         
         churn_dummy = case_when(
           default_tag !=0 ~ 1, 
           TRUE ~ 0
         )) %>%
  filter(FH_bill_issuetag !=1) %>%
  # create logarithmic transformation of kwh allocation for model later
  mutate(
    log_solar = case_when(
      kwh_allocated>0 ~ log(kwh_allocated),
      TRUE ~ 0
    )
  ) 
```




```{r message=FALSE, warning=FALSE, include=FALSE}

#group by the utility account number to get descriptive stats

flipped <- raw %>%
  group_by(utility_acct_number) %>%
  summarise(
            tenure = length(unique(date_concat)),
            default = sum(default_tag=="Defaulted Payment"),
            HH_Income = unique(`HH Income`),
            payment_method = unique(payment_type),
            temp_solar = max(kwh_allocated/1229),
            Churn = max(churn_dummy),
            solar_farm = unique(solar_farm_name)) %>%
            # Proj_Frog_Hollow_dummy = sum(solar_farm_name=="Frog Hollow (CHGE)"),
            # Proj_Howell_dummy = sum(solar_farm_name=="Howell (O&R)")) %>%
  mutate(defaulted = case_when(
    default>0 ~ 1,
    TRUE ~ 0
  ), count_user_id = n(),
  Income = case_when(
    HH_Income==">=$50,000" ~ "High",
    HH_Income=="$0-$49,999" ~ "Low"
  )) %>%
dplyr::select(-default, -HH_Income)


## filter out the duplicates


dupes <- flipped %>%
  filter(count_user_id == 2) %>%
  # duplicates are removed by only chosing the ACH payment method, there were duplicate 'card' payment entries 
  filter(payment_method == "ach")

no_dupes <- flipped %>%
  filter(count_user_id==1) 

```





```{r echo=FALSE, message=FALSE, warning=FALSE}
## join back with the main data

joined <- rbind(no_dupes, dupes) 
  
```


The first ten rows of the data are presented, grouped by account number. Unique identifiers were replaced with sequential IDs. 

* Tenure: length in months each account was active 
* Payment_Method: either card or ACH/direct debit method
* temp_solar: annual (right?) solar allocation in kWh 
* Churn: binary _var_iable capturing if an account left the solar farm
* defauted: binary variable capturing if account holder defaulted on payment
* Income: Low (<$50,000) or High (>$50,000)
* Solar_Farm: Either Howell or Frog Hollow


```{r echo=FALSE, message=FALSE, warning=FALSE}

##remove account ID and replace with dummy sequence
joined$Account_ID <- seq(1,nrow(joined),1)

temp <- joined[,-1]

temp <- temp %>%
  relocate(Account_ID)



head(temp, 10) %>%
  kbl(caption = "First 10 Home Rows",  digits=2) %>%
  kable_classic(full_width = T, html_font = "Cambria")
```

  




# Descriptive Statistics

## Total






The summary statistics below provide additional information. The average tenure of the 812 accounts was 20.3 months. The churn rate was 12.4%. The sample does skew towards high income homes, however note a large number of missing values were observed in income (567, 70%). 


```{r echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)

tenure <- joined %>%
  mutate(Name = "Tenure") %>%
  group_by(Name) %>%
  summarise(
    Total = sum(tenure),
    Mean = mean(tenure),
    Max = max(tenure),
    Min = min(tenure)
    
  )

defaults <- joined %>%
  mutate(Name = "Defaults") %>%
  group_by(Name) %>%
  summarise(
    Total = sum(defaulted, na.rm=T),
    Mean = mean(defaulted, na.rm=T),
    Max = max(defaulted, na.rm=T),
    Min = min(defaulted, na.rm=T)
    
  )

churn <- joined %>%
  mutate(Name = "Churn") %>%
  group_by(Name) %>%
  summarise(
    Total = sum(Churn, na.rm=T),
    Mean = mean(Churn, na.rm=T),
    Max = max(Churn, na.rm=T),
    Min = min(Churn, na.rm=T)
    
  )



low <- joined %>%
  mutate(Name = "Low Income") %>%
  group_by(Name) %>%
  summarise(
    Total = sum(Income=="Low", na.rm=T),
    Mean = mean(Income=="Low", na.rm=T),
    Max = NA,
    Min = NA
    
  )

high <- joined %>%
  mutate(Name = "High Income") %>%
  group_by(Name) %>%
  summarise(
    Total = sum(Income=="High", na.rm=T),
    Mean = mean(Income=="High", na.rm=T),
    Max = NA,
    Min = NA
    
  )

solar <- joined %>%
  mutate(Name = "Solar") %>%
  group_by(Name) %>%
  summarise(
    Total = sum(temp_solar, na.rm=T),
    Mean = mean(temp_solar, na.rm=T),
    Max =  max(temp_solar, na.rm=T),
    Min =  min(temp_solar, na.rm=T)
    
  )

payment_card <- joined %>%
  mutate(Name = "Payment by Card") %>%
  group_by(Name) %>%
  summarise(
    Total = sum(payment_method=="card", na.rm=T),
    Mean = mean(payment_method=="card", na.rm=T),
    Max =  NA,
    Min =  NA
    
  )

payment_ach <- joined %>%
  mutate(Name = "Payment by ACH") %>%
  group_by(Name) %>%
  summarise(
    Total = sum(payment_method=="ach", na.rm=T),
    Mean = mean(payment_method=="ach", na.rm=T),
    Max =  NA,
    Min =  NA
    
  )



summary_stats <- rbind(tenure,defaults,churn, low, high, solar, payment_card, payment_ach)



summary_stats %>%
  remove_rownames() %>%
  column_to_rownames(var = 'Name') %>%
  kbl(caption="Descriptive Statistics",  digits=2) %>%
  kable_classic(full_width = T, html_font = "Cambria")



```

## Frog Hollow





```{r echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)

temp <- joined %>%
  filter(solar_farm == "Frog Hollow (CHGE)")

tenure <- temp %>%
  mutate(Name = "Tenure") %>%
  group_by(Name) %>%
  summarise(
    Total = sum(tenure),
    Mean = mean(tenure),
    Max = max(tenure),
    Min = min(tenure)
    
  )

defaults <- temp %>%
  mutate(Name = "Defaults") %>%
  group_by(Name) %>%
  summarise(
    Total = sum(defaulted, na.rm=T),
    Mean = mean(defaulted, na.rm=T),
    Max = max(defaulted, na.rm=T),
    Min = min(defaulted, na.rm=T)
    
  )

churn <- temp %>%
  mutate(Name = "Churn") %>%
  group_by(Name) %>%
  summarise(
    Total = sum(Churn, na.rm=T),
    Mean = mean(Churn, na.rm=T),
    Max = max(Churn, na.rm=T),
    Min = min(Churn, na.rm=T)
    
  )



low <- temp %>%
  mutate(Name = "Low Income") %>%
  group_by(Name) %>%
  summarise(
    Total = sum(Income=="Low", na.rm=T),
    Mean = mean(Income=="Low", na.rm=T),
    Max = NA,
    Min = NA
    
  )

high <- temp %>%
  mutate(Name = "High Income") %>%
  group_by(Name) %>%
  summarise(
    Total = sum(Income=="High", na.rm=T),
    Mean = mean(Income=="High", na.rm=T),
    Max = NA,
    Min = NA
    
  )

solar <- temp %>%
  mutate(Name = "Solar") %>%
  group_by(Name) %>%
  summarise(
    Total = sum(temp_solar, na.rm=T),
    Mean = mean(temp_solar, na.rm=T),
    Max =  max(temp_solar, na.rm=T),
    Min =  min(temp_solar, na.rm=T)
    
  )

payment_card <- temp %>%
  mutate(Name = "Payment by Card") %>%
  group_by(Name) %>%
  summarise(
    Total = sum(payment_method=="card", na.rm=T),
    Mean = mean(payment_method=="card", na.rm=T),
    Max =  NA,
    Min =  NA
    
  )

payment_ach <- temp %>%
  mutate(Name = "Payment by ACH") %>%
  group_by(Name) %>%
  summarise(
    Total = sum(payment_method=="ach", na.rm=T),
    Mean = mean(payment_method=="ach", na.rm=T),
    Max =  NA,
    Min =  NA
    
  )



summary_stats <- rbind(tenure,defaults,churn, low, high, solar, payment_card, payment_ach)



summary_stats %>%
  remove_rownames() %>%
  column_to_rownames(var = 'Name') %>%
  kbl(caption="Frog Hollow Descriptive Statistics",  digits=2) %>%
  kable_classic(full_width = T, html_font = "Cambria")



```

## Howell




```{r echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)

temp <- joined %>%
  filter(solar_farm == "Howell (O&R)")

tenure <- temp %>%
  mutate(Name = "Tenure") %>%
  group_by(Name) %>%
  summarise(
    Total = sum(tenure),
    Mean = mean(tenure),
    Max = max(tenure),
    Min = min(tenure)
    
  )

defaults <- temp %>%
  mutate(Name = "Defaults") %>%
  group_by(Name) %>%
  summarise(
    Total = sum(defaulted, na.rm=T),
    Mean = mean(defaulted, na.rm=T),
    Max = max(defaulted, na.rm=T),
    Min = min(defaulted, na.rm=T)
    
  )

churn <- temp %>%
  mutate(Name = "Churn") %>%
  group_by(Name) %>%
  summarise(
    Total = sum(Churn, na.rm=T),
    Mean = mean(Churn, na.rm=T),
    Max = max(Churn, na.rm=T),
    Min = min(Churn, na.rm=T)
    
  )



low <- temp %>%
  mutate(Name = "Low Income") %>%
  group_by(Name) %>%
  summarise(
    Total = sum(Income=="Low", na.rm=T),
    Mean = mean(Income=="Low", na.rm=T),
    Max = NA,
    Min = NA
    
  )

high <- temp %>%
  mutate(Name = "High Income") %>%
  group_by(Name) %>%
  summarise(
    Total = sum(Income=="High", na.rm=T),
    Mean = mean(Income=="High", na.rm=T),
    Max = NA,
    Min = NA
    
  )

solar <- temp %>%
  mutate(Name = "Solar") %>%
  group_by(Name) %>%
  summarise(
    Total = sum(temp_solar, na.rm=T),
    Mean = mean(temp_solar, na.rm=T),
    Max =  max(temp_solar, na.rm=T),
    Min =  min(temp_solar, na.rm=T)
    
  )

payment_card <- temp %>%
  mutate(Name = "Payment by Card") %>%
  group_by(Name) %>%
  summarise(
    Total = sum(payment_method=="card", na.rm=T),
    Mean = mean(payment_method=="card", na.rm=T),
    Max =  NA,
    Min =  NA
    
  )

payment_ach <- temp %>%
  mutate(Name = "Payment by ACH") %>%
  group_by(Name) %>%
  summarise(
    Total = sum(payment_method=="ach", na.rm=T),
    Mean = mean(payment_method=="ach", na.rm=T),
    Max =  NA,
    Min =  NA
    
  )



summary_stats <- rbind(tenure,defaults,churn, low, high, solar, payment_card, payment_ach)



summary_stats %>%
  remove_rownames() %>%
  column_to_rownames(var = 'Name') %>%
  kbl(caption="Howell Descriptive Statistics",  digits=2) %>%
  kable_classic(full_width = T, html_font = "Cambria")



```



```{r message=FALSE, warning=FALSE, include=FALSE}

# Percentage of homes with any Default

any_defaults = sum(joined$defaulted, na.rm=T)/nrow(joined)

any_defaults 

any_churn = sum(joined$Churn>0, na.rm=T)/nrow(joined)


any_churn 
```

## Tenure Length


Obersvations are next grouped by tenure length. 






```{r echo=FALSE, message=FALSE, warning=FALSE}

tenures <- joined %>%
  mutate( tenure_length = case_when(
    tenure < 6 ~ "≤ 5",
    tenure > 5 & tenure < 11 ~ "5 to 10",
    tenure > 10 & tenure < 15 ~ "10 to 15",
    tenure >= 15~ "15 to 20",
    
  )) %>%
    group_by(tenure_length) %>%
    summarise(`Number of homes` = n(),
            `Churned` = sum(Churn, na.rm=T),
            `Defaulted` = sum(defaulted, na.rm=T),
            `Average Solar Allocation` = mean(temp_solar, na.rm=T),
            `Payment by Card` = sum(payment_method=="card", na.rm=T),
            `Payment by ACH` = sum(payment_method=="ach", na.rm=T)

            ) %>%
  mutate(
    temp_tag = case_when(
      tenure_length == "≤ 5" ~ 1,
      tenure_length == "5 to 10" ~ 2,
      tenure_length == "10 to 15" ~ 3,
      tenure_length == "15 to 20" ~ 4
    )
  ) %>%
  arrange(temp_tag) %>%
  dplyr::select(-temp_tag) %>%
  adorn_totals('row') 
    


tenures$`Average Solar Allocation`[tenures$tenure_length=="Total"] <- mean(joined$temp_solar,na.rm = T)

tenures %>%
 remove_rownames() %>%
  column_to_rownames(var = 'tenure_length') %>%
  kbl(caption="Account Tenure Length Statistics",  digits=2, format = "html", table.attr = "style='width:30%;'") %>%
  kableExtra::kable_styling() %>%
  kable_classic(full_width = F, html_font = "Cambria")



```


For homes that churned, the tenure was reasonbly much shorter than those that did not churn. The below graph shows the average tenure between these two groups, showing that it is almost double for those homes that exhibit continuous subscription. 

```{r echo=FALSE, message=FALSE, warning=FALSE}
temp <- joined %>%
  mutate(`Churn Status` = case_when(
    Churn == 0 ~ "Continuous",
    Churn ==1 ~ "Churned"
  )) %>%
  group_by(`Churn Status`) %>%
  summarise(`Average Months` = mean(tenure, na.rm=T)) %>%
  filter(!is.na(`Churn Status`))

p <- ggplot(data = temp, aes(x=factor(`Churn Status`), y = `Average Months`)) + geom_bar(stat='identity') + theme_minimal() + xlab("") +ggtitle("Average Tenure by Churn")


ggplotly(p)


temp <- joined %>%
  mutate(`Default Status` = case_when(
    defaulted == 0 ~ "Non-Default",
    defaulted > 0 ~ "Default"
  )) %>%
  group_by(`Default Status`) %>%
  summarise(`Average Months` = mean(tenure, na.rm=T))%>%
  filter(!is.na(`Default Status`))

p <- ggplot(data = temp, aes(x=factor(`Default Status`), y = `Average Months`)) + geom_bar(stat='identity') + theme_minimal() + xlab("")+ggtitle("Average Tenure by Default")

ggplotly(p)

```



## Income


```{r echo=FALSE, message=FALSE, warning=FALSE}

library(kableExtra)
library(janitor)

# group_by_home$dummy <- 1
# 
# summary_table <- group_by_home %>%
#   group_by(dummy) %>%
#   summarise(`Number of homes` = n(),
#             `Low Income` = sum(Income=="Low", na.rm=T),
#             `High Income` = sum(Income=="High", na.rm=T),
#             `Average Days Late/Month` = mean(Avg_DaysLate),
#             `Average Solar Allocation` = mean(Avg_Solar, na.rm=T)) %>%
#   select(-dummy)
# 
# 
# 
# summary_table %>%
#   kbl(caption = "cenhud Churn & Default Descriptive Statistics",  digits=2) %>%
#   kable_classic(full_width = F, html_font = "Cambria")


summary_table <- joined %>%
  group_by(Income) %>%
  summarise(`Number of accounts` = n(),
            `Churned` = sum(Churn, na.rm=T),
            `Defaulted` = sum(defaulted, na.rm=T),
            `Average Solar Allocation` = mean(temp_solar, na.rm=T),
            `Payment by Card` = sum(payment_method=="card", na.rm=T),
            `Payment by ACH` = sum(payment_method=="ach", na.rm=T)
            ) %>%
  adorn_totals('row') 

summary_table$`Average Solar Allocation`[summary_table$Income=="Total"] <- mean(joined$temp_solar,na.rm = T)


summary_table %>%
  kbl(caption = "Resident Income Statistics",  digits=2) %>%
  kable_classic(full_width = F, html_font = "Cambria")
```


## Churn Reasons

When a respondent left the program, a reason was recorded, when available. The below summarises the accounts that experienced a churn, which may further be collapsed into categories. Note the total below may not match the churn totals in the above descriptive statistics due to defaulted payments and duplicates included. 

```{r eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}

# churn_reasons <- joined %>%
#   group_by(`Churn Reasons`= churn_reason) %>%
#  # filter(churn_reason !=0) %>%
#   summarise(Total=n(),
#             `High Income` = sum(Income=="High", na.rm=T),
#             `Low Income` = sum(Income=="Low", na.rm=T),
#             `No Income` = sum(is.na(Income))) %>%
#   arrange(-Total) %>%
#   adorn_totals('row')
# 
# churn_reasons%>%
#   kbl() %>%
#   kable_classic(full_width = F, html_font = "Cambria")

```



```{r eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}
#unit of analysis work:

#Group by Homes
home_groups <- raw %>%
  group_by(property_id) %>%
  summarise(total=n(),
            accounts = sum(!is.na(utility_acct_number)),
            users = sum(!is.na(user_id)),
            accounts_churn = sum(default_tag != 0 & default_tag != "Defaulted Payment"))

home_groups
sapply(home_groups, function(x) sum(is.na(x)))

#Group by Users
user_groups <- raw %>%
  group_by(user_id) %>%
  summarise(total=n(),
            accounts = sum(!is.na(utility_acct_number)),
            homes = sum(!is.na(property_id)))

user_groups
sapply(user_groups, function(x) sum(is.na(x)))

#Group by Account
account_groups <- raw %>%
  group_by(utility_acct_number) %>%
  summarise(total=n(),
           users = sum(!is.na(user_id)),
            homes = sum(!is.na(property_id)))

account_groups
sapply(account_groups, function(x) sum(is.na(x)))

```



```{r}
temp <- raw %>%
 group_by(user_id,utility_acct_number) %>%
  summarise(tenure = length(unique(date_concat)),
            count_accts = n(),
            default = sum(default_tag=="Defaulted Payment"),
           # count_accts = length(unique(utility_acct_number)) ,
            count_churn = sum(default_tag !=0 & default_tag !="Defaulted Payment")/length(unique(date_concat))) #%>%
 # filter(count_accts > 1) 

temp

```









# Results









## Churn Figures 

### Solar

```{r echo=FALSE, message=FALSE, warning=FALSE}
library(ggplot2)
library(plotly)

myColors <- c("green", "red")

# p <- ggplot(data=group_by_home, aes(x=Income, y = Avg_DaysLate, color=factor(Default_Tag))) + geom_point() +  scale_color_manual(values=myColors, labels=c('No', 'Yes')) +
#   labs(colour="Defaulted") + theme_classic()
# 
# ggplotly(p)


p <- ggplot(data=joined, aes(x=Income, y = temp_solar, color=factor(Churn))) + geom_point() +  scale_color_manual(values=myColors, labels=c('No', 'Yes')) +  labs(colour="Churn Status") + ylab("Solar kWh") +theme_classic() 

ggplotly(p)


p <- ggplot(data=joined, aes(x=payment_method, y = temp_solar, color=factor(Churn))) + geom_point() +  scale_color_manual(values=myColors, labels=c('No', 'Yes')) +  labs(colour="Churn Status") + ylab("Solar kWh") +theme_classic() 

ggplotly(p)

p <- ggplot(data=joined, aes(x=tenure, y = temp_solar, color=factor(Churn))) + geom_point() +  scale_color_manual(values=myColors, labels=c('No', 'Yes')) +  labs(colour="Churn Status") + ylab("Solar kWh") +theme_classic() + geom_smooth(data=subset(joined,Churn==1 | Churn==0),
               aes(tenure,temp_solar,color=factor(Churn)),method=lm,se=FALSE)

ggplotly(p)

```




### Tenure & Churn by Payment Method 

```{r echo=FALSE, message=FALSE, warning=FALSE}



p <- ggplot(data=joined, aes(x=Income, y = tenure, color=factor(Churn))) + geom_point(stat="identity") +  scale_color_manual(values=myColors, labels=c('No', 'Yes')) +  labs(colour="Churn Status") + ylab("Tenure") +theme_classic() 

ggplotly(p)

p <- ggplot(data=joined, aes(x=Income, y = tenure, color=factor(Churn))) + geom_bar(stat="identity") +  scale_color_manual(values=myColors, labels=c('No', 'Yes')) +  labs(colour="Churn Status") + ylab("Tenure") +theme_classic() 

ggplotly(p)


p <- ggplot(data=joined, aes(x=payment_method, y = tenure, color=factor(Churn))) + geom_point(stat="identity") +  scale_color_manual(values=myColors) +  labs(colour="Churn Status") + ylab("Tenure") +theme_classic() 

ggplotly(p)


p <- ggplot(data=joined, aes(x=payment_method, y = tenure, color=factor(Churn))) + geom_bar(stat="identity") +  scale_color_manual(values=myColors) +  labs(colour="Churn Status") + ylab("Tenure") +theme_classic() 

ggplotly(p)
```







## Default Figures 

### Solar

```{r echo=FALSE, message=FALSE, warning=FALSE}
library(ggplot2)
library(plotly)

myColors <- c("green", "red")

# p <- ggplot(data=group_by_home, aes(x=Income, y = Avg_DaysLate, color=factor(Default_Tag))) + geom_point() +  scale_color_manual(values=myColors, labels=c('No', 'Yes')) +
#   labs(colour="Defaulted") + theme_classic()
# 
# ggplotly(p)


p <- ggplot(data=joined, aes(x=Income, y = temp_solar, color=factor(defaulted))) + geom_point() +  scale_color_manual(values=myColors, labels=c('No', 'Yes')) +  labs(colour="Churn Status") + ylab("Solar kWh") +theme_classic() 

ggplotly(p)


p <- ggplot(data=joined, aes(x=payment_method, y = temp_solar, color=factor(defaulted))) + geom_point() +  scale_color_manual(values=myColors, labels=c('No', 'Yes')) +  labs(colour="Churn Status") + ylab("Solar kWh") +theme_classic() 

ggplotly(p)

p <- ggplot(data=joined, aes(x=tenure, y = temp_solar, color=factor(defaulted))) + geom_point() +  scale_color_manual(values=myColors, labels=c('No', 'Yes')) +  labs(colour="Churn Status") + ylab("Solar kWh") +theme_classic() + geom_smooth(data=subset(joined,defaulted==1 | defaulted==0),
               aes(tenure,temp_solar,color=factor(defaulted)),method=lm,se=FALSE)

ggplotly(p)

```




### Tenure & Default by Payment Method 

```{r echo=FALSE, message=FALSE, warning=FALSE}



p <- ggplot(data=joined, aes(x=Income, y = tenure, color=factor(defaulted))) + geom_point(stat="identity") +  scale_color_manual(values=myColors, labels=c('No', 'Yes')) +  labs(colour="Default Status") + ylab("Tenure") +theme_classic() 

ggplotly(p)

p <- ggplot(data=joined, aes(x=Income, y = tenure, color=factor(defaulted))) + geom_bar(stat="identity") +  scale_color_manual(values=myColors, labels=c('No', 'Yes')) +  labs(colour="Default Status") + ylab("Tenure") +theme_classic() 

ggplotly(p)


p <- ggplot(data=joined, aes(x=payment_method, y = tenure, color=factor(defaulted))) + geom_point(stat="identity") +  scale_color_manual(values=myColors) +  labs(colour="Default Status") + ylab("Tenure") +theme_classic() 

ggplotly(p)


p <- ggplot(data=joined, aes(x=payment_method, y = tenure, color=factor(defaulted))) + geom_bar(stat="identity") +  scale_color_manual(values=myColors) +  labs(colour="Default Status") + ylab("Tenure") +theme_classic() 

ggplotly(p)
```
















# Model 

A logit model is deployed to analyze the probability of a home churning. The model in brief is described below. Without additional data collected from Experian, limited variability can be measured due to NA values in addition to limited number of coefficients to measure. 

In the limited data available, statistically significant effects were measured for probability of churn for both tenure (Months) and days late. Due to relatively small number of observations of late payments, no statistically significant results were measured in probability of late or default payments. 


\begin{equation}

P_{d} = \beta_{0} + \beta_{1}*Months +\beta_{2}*(Payment Method = Card) +\beta_{3}*Income=Low \\ +\beta_{4}*Solar kwH + \beta{5}*(Solar Farm=Howell)

\end{equation}



```{r echo=FALSE, message=FALSE, warning=FALSE}

# Logit Models

joined$Income <- as.factor(joined$Income)


#model_default <- glm(Default_Tag ~ Months +  Avg_DaysLate + Avg_Solar + Income + DPD_30_Tag + DPD_60_Tag + DPD_90_Tag, family=binomial(link='logit'), data=group_by_home)


model_churn <- glm(Churn ~ tenure  + payment_method + Income + temp_solar + solar_farm, family=binomial(link='logit'), data=joined)
summary(model_churn)



```





The logit curves below show the relationship between the distribution of both tenure and days late with the probability of churning. The first graph shows that the longer the tenure, the lower the likelihood. The second graph shows that the days of late payments are generally not associated with likelihood of churn. 

```{r echo=FALSE, message=FALSE, warning=FALSE}

# Logit Curves - none significant except churn


ggplot(joined, aes(x=tenure, y = Churn))+ geom_point(fill='red') + 
  stat_smooth(method="glm", method.args=list(family=binomial(link='logit')), se=FALSE)+theme_minimal()+ scale_x_continuous(name="Months") +
  scale_y_continuous(name="Likelihood of Churn")




ggplot(joined, aes(x=temp_solar, y = Churn))+ geom_point(fill='red') + 
  stat_smooth(method="glm", method.args=list(family=binomial(link='logit')), se=FALSE)+theme_minimal()+ scale_x_continuous(name="Solar kWh") +
  scale_y_continuous(name="Likelihood of Churn")

```





# Discussion

