---
title: "Churn and Default"
author: "Jacob Ford, Solstice Data Scientist"
date: "`r Sys.Date()`"
output: 
  rmdformats::readthedown:
    toc_depth: 5

knit: (function(inputFile, encoding) { 
        rmarkdown::render(inputFile, encoding = encoding, 
        output_file = file.path(dirname(inputFile), 'churn_and_default.html')) })
---






# Introduction

For the final component of the research, we shift our attention to customer enrollment in current community solar farms. By analyzing tenure length and the presence of an individual departing community solar farms, our research seeks to provide quantitative performance data to add depth to the previously collected survey data. 

Given the incipient nature of community solar, this analysis provides an opportunity to quantify the rates of churn or default while controlling for household or demographic attributes. The primary goal of the analysis is to first describe the characteristics of the residents, the prevalence of default and/or churn rates, the prevalence of churn rates, and any statistically significant differences between groups in their default, late payments or churn rates. Secondarily, the analysis will seek to determine if enrollment in the community solar farm is associated with any measurable change in credit scores. 

Customers in existing community solar projects are evaluated for their subscription and payment status on a monthly basis. Participation in community solar farms is voluntary, hence customers may exit the program at any point, often subject to a cancellation fee set out in the contract. Additionally, customers sometimes fail to pay the monthly solar farm subscription fee. Churn refers to the act of a customer exiting the solar farm, whereas default refers to the customer failing to pay the monthly subscription fee.

# Methods

Customer data is collected from two community solar projects in New York.  Monthly account level data was collected from January 2020 to April 2022. Account level information includes monthly payment performance, for which churn and default triggers are captured. Monthly, account level performance data is appended using Experian data including Experian's Income Insight Score and Vantage Score, along with demographic data such as gender, education, occupation, and homeownership status. 

For the primary analysis, a logit model is used to determine likelihood of churn and/or default, taking into account an account's length of tenure and various demographic and socioeconomic data available. 

The secondary analysis is performed to measure the difference in credit scores for customers, to measure if any difference of statistical importance is observed over the length of the enrollment in the farm. Credit scores for each customer are measured in December 2020 and April 2022, offering an approximate pre- and post credit score measurement. A simple difference of means test (Welch) is employed to measure the differences. 



## Data Cleaning

Data from the solar farms consists of 32,384 monthly observations over 812 utility accounts and 620 unique users. Multiple accounts may be tied to a single user, if multiple residences are owned or if a resident churned and subsequently returned in the observed time period. 




```{r message=FALSE, warning=FALSE, include=FALSE}

#Load packages we'll need throughout the entire document

library(readxl)
library(sf)
library(dplyr)
library(janitor)
library(dplyr)
library(tidyverse)
library(nngeo)
library(plotly)
library(kableExtra)
library(haven)
library(rms)
library(rmdformats)
library(gtsummary)
library(ggplot2)
library(plotly)
library(gridExtra)
library(grid)
library(lattice)
library(stargazer)


options(scipen = 100)


```

```{r message=FALSE, warning=FALSE, include=FALSE}

## Change the below to match your local drive ##

input_folder <- '/Volumes/GoogleDrive-111165283491444964488/Shared drives/Solstice | Low Income Inclusion/05_Research/05.01_SETO/DOE SETO Project Implementation/Project Management/SETO Data Analysis/3. Churn and Default/Files for NMR Review/'

setwd(input_folder)

```



Data cleaning mainly involves removing observations for one farm over a particular time period. Data was collected from two sites, referred to as Farm A and Farm B for privacy. As a result of a utility billing issue wherein payment performance was not available starting in October 2021 and not recommencing until April 2022 at Farm A. Hence, the values for October, November and December 2021 are removed. This narrowed dataset consists of 31,704 observations; the number of utility accounts remained constant at 812 and the number of unique users stays constant at 620. 


```{r message=FALSE, warning=FALSE, include=FALSE}

# 'Nautilus Raw Data for NMR.xlsx' contains the monthly enrollment records for the two sample solar farms - contains various account specific information. We rename this 'farm_data'. For the analysis, we do not differentiate between the two farm locations, however for descriptive statistics we break out the differences for additional clarity. 

# There are four main ID variables in the 'farm_data' data frame: user (user_id),  household (property_id), utility account (utility_acct_number), and Participant ID (ParticipantID). ParticipantID is a mask of user_id and is used to merge the data with Experian attributes. We decided to group the analysis by utility accounts. The reason is we are measuring is rate of churn/default, and the account is the unit that performs these actions; not the user which may hold multiple accounts. Household ID and utility account are very similar, but for consistency, we hereafter focus on utility account data


farm_data <- read_excel(paste(input_folder, "Nautilus Raw Data for NMR.xlsx", sep=""), 
    sheet = "final data")  %>%
  # make tag for Farm A properties after 9/2021 ; this results in just dropping October 2021 values for Farm A
  mutate(FH_bill_issuetag = case_when(
         solar_farm_name=="Farm A" & year == 2021 & month > 9 ~ 1,
         TRUE ~ 0)) %>%
  filter(FH_bill_issuetag !=1)  %>%
  mutate(utility_acct_number = as.factor(utility_acct_number))
```


## Generate Churn and Default Tags

For the accounts that left the solar farm at any point, we generate Churn and Default 'triggers'. The table below shows the variety of reasons listed why a particular account left either solar farm during the observed period. Most left the solar farm as a result of moving from the service area or no longer interested in the solar farm. 

Defaulted accounts are triggered when the leave reason is 'Defaulted Payment'; every other leave reason is marked as churn, capturing a wider array of outcomes, from moving to health reasons. Table 1 shows how reasons listed for an account leaving a farm were mapped to Churn and Default triggers. Note, we categorize every default as a churn event as well for simplicity, however not every churn event is a default. 

#### Table 1: Churn and Default Triggers


```{r echo=FALSE, message=FALSE, warning=FALSE}



#  'leave_reason' - trigger that captures reason why an individual leaves a solar farm; this variable is used to create our dependent variables, 
# 'Churn' and 'Default'

farm_data %>%
  filter(leave_reason!=0) %>%
  group_by(Reason = leave_reason) %>%
  summarise(N=n()) %>%
  mutate(Trigger = case_when(
    grepl("Default", Reason) ~ "Default",
    TRUE ~ "Churn"
  )) %>%
  arrange(desc(N)) %>%
  select(-N) %>%
    kbl(caption = "Table 1: Churn and Default Triggers",
      digits = 3, format.args = list(big.mark = ",",scientific = FALSE)) %>%
  kable_classic(full_width = T, html_font = "Cambria") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) 


# Create churn/default dummy variables

farm_data <- farm_data %>%
  mutate(
# Create dummy variable for default, the more extreme case of churn; every default is a churn, not ever churn is a default of course
         default_dummy = case_when(
           leave_reason =="Defaulted Payment" ~ 1,
           TRUE ~ 0),
         churn_dummy = case_when(
           leave_reason !=0 ~ 1, 
           TRUE ~ 0)
  )

```



## Filtering Payment Methods

For some accounts, both a credit card and direct deposit (ACH) payment methods were recorded for the same month. To remove these entries, we first flip the dataframe by utility account ID, then isolate the duplicates and only take the payment method for direct deposit, however using credit card payment as the method would result in the same outcome, as the underlying performance or attribute data does not differ. 


```{r message=FALSE, warning=FALSE, include=FALSE}

#group by the utility account number to get descriptive stats

# note we use "unique" here for string variables and "max" for continuous variables. This is done for convenience, as these variables (besides payment_method described below) do not vary, hence we are just tagging the utility accounts it's corresponding leave reason, churn and default tag, etc. 

flipped <- farm_data %>%
  group_by(utility_acct_number) %>%
  summarise(
            n= n(),
            ParticipantID = max(ParticipantID),
            tenure = length(unique(date_concat)), #measures number of months each utility account is observed in the 'farm_data' df
            payment_method = unique(payment_type),
            leave_reason = unique(leave_reason),
            
            # divide kwh_allocated by 1229 to get kWh per account level, This number was generated internally by Solstice and is specific to the projects managed
            kwh_solar = mean(kwh_allocated/1229),
            Churn = max(churn_dummy),
            Default = max(default_dummy),
            solar_farm = unique(solar_farm_name)
            ) %>%
  mutate(
    count_user_id = n()) 

## filter out the duplicates


dupes <- flipped %>%
  filter(count_user_id == 2) %>%
  # duplicates are removed by only chosing the ACH payment method, there were duplicate 'card' payment entries 
  filter(payment_method == "ach")

no_dupes <- flipped %>%
  filter(count_user_id==1) 

```





```{r echo=FALSE, message=FALSE, warning=FALSE}
## join back with the main data

joined <- rbind(no_dupes, dupes) %>%
  select(-n,-count_user_id) 


```





## Append Experian Data

The data has been grouped by utility account IDs. As seen in Table 2, the information available as collected by Solstice includes:

- utility account number
- Participant ID
- Tenure (in months) of account
- payment method
- leave_reason, as described above
- kWh allotted to each utility account (note this figure is arrived at by dividing the raw kWh reported by 1,229: this is a project specific conversion rate)
- Churn and Default dummy variables
- solar farm identification

#### Table 2: Sample of Data Collected by Solstice

```{r echo=FALSE, message=FALSE, warning=FALSE}
joined[200:210, ] %>%
    kbl(caption = "Table 2: Sample of Data Collected by Solstice",
      digits = 3, format.args = list(big.mark = ",",scientific = FALSE)) %>%
  kable_classic(full_width = T, html_font = "Cambria") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) 
```

To add additional data to the analysis, we contracted with Experian to provide additional demographic and socioeconomic data. This data was provided on the *Participant* level. Recall that multiple utility accounts may be held by a single participant. The data provided by Experian has a matching 'ParticipantID' tag, that corresponds to the Solstice data set. 

Note: Not all Participants and Utility accounts are able to be appended, due to limited data availability from Experian. The 'pin rate' refers to the rate at which data is successfully collected for a record. For the 620 Participant_Ids, 506 received data, a pin rate of 81.6%. However, even succesfully pinned records did not receive full data, or possibly received an 'unknown' response. These differences are observed in Tables 4 and 5. 

Data from Experian was provided for two different time periods, December 2020 and April 2022. The utility of having differing time periods is to allow for the secondary analysis, wherein change in credit scores is calculated and a difference of means test is conducted to determine if tenure in the solar farms is associated with any measurable change in credit score. An average credit score variable is created and used in the primary analysis as well. 

For the primary analysis, all other non-credit score Experian data is appended to the Solstice data set at the December 2020 time frame. The December 2020 data is chosen for two reasons: firstly, little change is observed or expected to be observed in demographic data such as homeownership status, education, marriage, and gender; hence there is little difference in terms of model construction between using one or the other source. Secondly, and more importantly, the data provided as of December 2020 is slightly more complete, meaning fewer data are missing for fewer participants.  

Finally, Solstice included additional non-SETO data to be scored by Experian to satisfy Experian minimum requirement requirements. It is essential to remove this data from the analysis. Participant IDs from 1-621 refer to SETO related entries; values > 621 are filtered out of the analysis. 

Table 3 below shows a sample of the appended data. Note that VANTAGE_V4_SCORE.x refers to the Vantage Score as of April 2022, and VANTAGE_V4_SCORE.y refers to Vantage Scores as of December 2020. 



```{r echo=FALSE, message=FALSE, warning=FALSE}


## Merge with Experian Data, start with demographic

score_april_22 <- read.csv(paste(input_folder, 'From Experian/Data/202223148A_043022_SCORE_DEMO.csv', sep=""))
score_april_22$period <- "April 2022"

score_dec_20 <- read.csv(paste(input_folder, 'From Experian/Data/202223148A_123020_SCORE_DEMO.csv', sep=""))
score_dec_20$period <- "December 2020"


total_score <- merge(score_april_22,score_dec_20, by="PARTICIPANT_ID") %>%
  filter(PARTICIPANT_ID <= 621) %>% #filter non-SETO Participants Out
  group_by(ParticipantID=PARTICIPANT_ID) %>%
  summarise(
            VANTAGE_V4_SCORE.x = VANTAGE_V4_SCORE.x, # .x refers to April 2022
            VANTAGE_V4_SCORE.y = VANTAGE_V4_SCORE.y, # .y refers to December 2020
            
            Vantage_Diff = VANTAGE_V4_SCORE.x - VANTAGE_V4_SCORE.y, # This measures the change in Vantage Score from Dec'20 -> Apr '22, 
                                                                    # positive number refers to positive change
                                                                    # observed over this time period; whereas a negative                                                                                               # number refers to a decline in Vantage Score.
            
            #Experian Demographic Variables, definitions for each are given at bottom of this chunk
            INCOME_INSIGHT_SCORE = INCOME_INSIGHT_SCORE.y,
            GENDER_CODE.y,
            MARRIAGE_IN.y,
            OCCUPATION_CD.y,
            HOMEOWNER_IN.y, 
            RENTER_IN.y,
            INDV_EDUCATION_CD.y
            ) %>%
  mutate(
      Vantage_Avg = (VANTAGE_V4_SCORE.x + VANTAGE_V4_SCORE.y)/2,
      GENDER = case_when(
        GENDER_CODE.y == "F" ~ "Female", 
        GENDER_CODE.y == "M" ~ "Male"),
     
       MARRIAGE = case_when(
        MARRIAGE_IN.y == "S" ~ "Single", 
        MARRIAGE_IN.y == "M" ~ "Married",
        MARRIAGE_IN.y == "U" ~ "Unknown"),
      
       HOMEOWNER = case_when(
        HOMEOWNER_IN.y == "Y" ~ "Yes", 
        HOMEOWNER_IN.y == "U" ~ "Unknown"),
      
      RENTER = case_when(
        RENTER_IN.y == "Y" ~ "Yes", 
        RENTER_IN.y == "U" ~ "Unknown"),
      
      EDUCATION = case_when(
        INDV_EDUCATION_CD.y %in% c(11,51,12,52)  ~ "High School or Some College",
        INDV_EDUCATION_CD.y %in% c(13,53)  ~ "Completed College",
        INDV_EDUCATION_CD.y %in% c(14,54)  ~ "Graduate Degree",
        INDV_EDUCATION_CD.y %in% c(15, 55, 0)  ~ "Other"),
      
      OCCUPATION = case_when(
        OCCUPATION_CD.y %in% c("06", "10", "23", "53","30", "32", "33", "34", "54",
                               "05", "18", "22", "31", "37") ~ "Services",
        OCCUPATION_CD.y %in% c("17", "21", "39", "40", "41", "42",
                               "43", "44", "45", "46", "47", "48",
                               "49", "50", "52") ~ "Healthcare/Education Services",
        OCCUPATION_CD.y %in% c("02", "03", "04", "13", "14",
                               "36", "38", "51") ~ "Management/Technical",
        OCCUPATION_CD.y %in% c("U", "00", "08", "11", "16") ~ "Self Employed/Other",
        OCCUPATION_CD.y %in% c("07", "12", "15", "35") ~ "Trade/Agriculture")) %>%
  select(-GENDER_CODE.y, -MARRIAGE_IN.y, -HOMEOWNER_IN.y, -RENTER_IN.y, -OCCUPATION_CD.y, -INDV_EDUCATION_CD.y)

# Replace INCOME_INSIGHT_SCORE = 9001, which refers to an income 'decrease' - see definition below in this chunk for more information. 

total_score$INCOME_INSIGHT_SCORE[total_score$INCOME_INSIGHT_SCORE==9001] <- NA

final_joined <- merge(joined, total_score , by="ParticipantID", all.x = TRUE)


## --- Occupation Job Codes Collapse --- ## 

# Categories	New_Categories
# 00=Unknown	Self Employed/Other
# 02=Professional/Technical	Management/Technical
# 03=Upper Management/Executive	Management/Technical
# 04=Middle Management	Management/Technical
# 05=Sales/Marketing	Services
# 06=Clerical/Office	Services
# 07=SkilledTrade/Machine/Laborer	Trade/Agriculture
# 08=Retired	Self Employed/Other
# 10=Executive/Administrator	Services
# 11=Self Employed	Self Employed/Other
# 12=Professional Driver	Trade/Agriculture
# 13=Military	Management/Technical
# 14=Civil Servant	Management/Technical
# 15=Farming/Agriculture	Trade/Agriculture
# 16=Work From Home	Self Employed/Other
# 17=Health Services	Healthcare/Education Services
# 18=Financial Services	Services
# 21=Teacher/Educator	Healthcare/Education Services
# 22=Retail Sales	Services
# 23=Computer Professional	Services
# 30=Beauty	Services
# 31=Real Estate	Services
# 32=Architects	Services
# 33=Interior Designers	Services
# 34=Landscape Architects	Services
# 35=Electricians	Trade/Agriculture
# 36=Engineers	Management/Technical
# 37=Accountants/CPA	Services
# 38=Attorneys	Management/Technical
# 39=Social Worker	Healthcare/Education Services
# 40=Counselors	Healthcare/Education Services
# 41=Occupational Ther/ Physical Ther	Healthcare/Education Services
# 42=Speech Path./Audiologist	Healthcare/Education Services
# 43=Psychologist	Healthcare/Education Services
# 44=Pharmacist	Healthcare/Education Services
# 45=Opticians/Optometrist	Healthcare/Education Services
# 46=Veterinarian	Healthcare/Education Services
# 47=Dentist/Dental Hygienist	Healthcare/Education Services
# 48=Nurse	Healthcare/Education Services
# 49=Doctors/Physicians/Surgeons	Healthcare/Education Services
# 50=Chiropractors	Healthcare/Education Services
# 51=Surveyors	Management/Technical
# 52=Clergy	Healthcare/Education Services
# 53=Insurance/Underwriters	Services
# 54=Services/Creative	Services



## --- Income Insight Score--- ## 

# INCOME_INSIGHT_SCORE
# 
# 0015 - 0990
# Score represents income in thousands of dollars, e.g.. 0100 = $100,000
# 
# Model Exclusions:
# 9001 = deceased
# 9002 - no update in 6 months
# 9003 = lack of credit history



## --- Education INDV_EDUCATION_CD--- ##

# First byte = model confidence: 1 = Extremely likely; 5 = Likely.
# We combine the first byte 1 and 5 values, Extremely likely + Likely 
# 
# Second Byte = education value: 
# 0=Unknown
# 1=High School Diploma,
# 2=Some College
# 3=Completed College
# 4=Graduate Degree
# 5=Retired
# blank=Blank


## --- Marriage MARRIAGE_IN--- ##

# M = Married
# S = Single
# U = Unknown

 
## --- Homeownership HOMEOWNER_IN--- ##

# Y = YES
# U = Unknown	


## --- Renter RENTER_IN--- ##

# Y = YES
# U = Unknown	




# Convert Experian Demographic Data into factors

final_joined$EDUCATION <- as.factor(final_joined$EDUCATION)


```


#### Table 3: Sample of Appended Data


```{r echo=FALSE, message=FALSE, warning=FALSE}
head(final_joined) %>%
    kbl(caption = "Table 3: Sample of Appended Data",
      digits = 3, format.args = list(big.mark = ",",scientific = FALSE)) %>%
  kable_classic(full_width = T, html_font = "Cambria") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) 
```


```{r message=FALSE, warning=FALSE, include=FALSE}
# The following code creates a dataframe "all_data" that adds the additional financial trade data provided by Experian. The 'premiers glossary' excel sheet has definitions for each of the premier variables. This data is not used in the analysis, just provided here for record. 

# Full Merge with Premiere Attributes

prem_attr_1 <- read.csv(paste(input_folder, 'From Experian/Data/202223148A_123020_PREMAB.csv', sep=""))
prem_attr_2 <- read.csv(paste(input_folder, 'From Experian/Data/202223148A_123020_PREMCD.csv', sep=""))

prem_attr<- merge(prem_attr_1, prem_attr_2 , by="PARTICIPANT_ID", all.y = TRUE) %>%
  mutate(ParticipantID=PARTICIPANT_ID) %>%
  select(-CTK.x, -CTK.y, -RECORD_NB.x,-RECORD_NB.y, -PARTICIPANT_ID)


all_data <- merge(final_joined, prem_attr , by="ParticipantID", all.x = TRUE) 


```



# Descriptive Statistics

Account level summary statistics by solar farm are provided in .  Note, only 6 instances of default were observed. However, of the 812 accounts, 118 or just over 14.5% of accounts, experienced churn. 

#### Table 4: Summary Statistics by Solar Farm


```{r echo=FALSE, message=FALSE, warning=FALSE}
temp <-
  all_data %>%
  ungroup() %>%
  select(Churn, Default,
         `Average Tenure` = tenure,  solar_farm, `Average kWh` = kwh_solar, `Payment Method`=payment_method,
         Gender = GENDER, `Marital Status` = MARRIAGE, Occupation = OCCUPATION, Education = EDUCATION, 
         Renter = RENTER, Homeowner = HOMEOWNER, INCOME_INSIGHT_SCORE,
         `Average Vantage Score` = Vantage_Avg,
         `Vantage Score April 2022` =VANTAGE_V4_SCORE.x, 
         `Vantage Score December 2020` =VANTAGE_V4_SCORE.y,`Vantage Score Change` = Vantage_Diff ) %>%
  tbl_summary(by = solar_farm,
              missing = "no") %>%
  modify_header(label = "**Table 4: Summary Statistics by Solar Farm**") %>% 
  bold_labels() %>%
  add_overall()

temp
```


Average tenure for the entire dataset is just under two years (22 months). Table 5 provides descriptive statistics grouped by tenure length. Most observations of Churn occurred within the 6-12 month range, 81% of all observed churn. 

#### Table 5: Summary Statistics by Tenure Length


```{r echo=FALSE, message=FALSE, warning=FALSE}
#joined$tenure_length <- factor(joined$tenure_length, levels=c("Less than 6", "6-12", "12-24", "Over 24"))


#levels(joined$tenure_length) <- c("Less than 6", "6-12", "12-24", "Over 24")

temp <-
  all_data %>%
    mutate( tenure_length = case_when(
    tenure < 6 ~ 0,
    tenure > 6 & tenure < 12 ~ 1,
    tenure > 12 & tenure < 24 ~ 2,
    tenure >= 24~ 3
  )) %>% 
  mutate(tenure_length_2 = recode(tenure_length,
         '0' = "< 6 Months",
         '1' = "6-12 Months",
         '2' = "12-24 Months",
         '3' = "Over 24 Months")) %>%
  
  arrange(desc(tenure_length)) %>%
  ungroup() %>%
  select(Churn, Default,
         `Average Tenure` = tenure, tenure_length_2, `Average kWh` = kwh_solar, `Payment Method`=payment_method,
         Gender = GENDER, `Marital Status` = MARRIAGE, Occupation = OCCUPATION, Education = EDUCATION, 
         Renter = RENTER, Homeowner = HOMEOWNER, INCOME_INSIGHT_SCORE,
         `Average Vantage Score` = Vantage_Avg,
         `Vantage Score April 2022` =VANTAGE_V4_SCORE.x, 
         `Vantage Score December 2020` =VANTAGE_V4_SCORE.y,`Vantage Score Change` = Vantage_Diff ) %>%
  tbl_summary(by = tenure_length_2, missing="no") %>%
  modify_header(label = "**Table 5: Summary Statistics by Tenure Length**") %>% 
  bold_labels() %>%
  add_overall()

temp
```








# Primary Analysis


## Model 

To determine the probability an account will experience either churn or default, logistic regression is used to analyze likelihood of either churn or default, using the demographic and socieconomic data available from primary data collected. The model employed is described in below: 

\begin{equation}

Churn_{Prob} = \beta_{0} + \beta_{1}Tenure + \beta_{2}kWh +\beta_{3}Gender \\ \beta_{4}VantageScore + \beta_{5}Education + \beta_{6}Married + \beta_{7}IncomeInsight

\end{equation}


\begin{equation}

Default_{Prob} = \beta_{0} + \beta_{1}Tenure + \beta_{2}kWh +\beta_{3}Gender \\ \beta_{4}VantageScore + \beta_{5}Education + \beta_{6}Married + \beta_{7}IncomeInsight

\end{equation}

Logarithmic transformations for continuous variables are employed. These include for Vantage Score and Income Insight Score. A number of models are considered, each taking different combinations of input data. The results are shown below in Table 6.  

```{r echo=FALSE, message=FALSE, warning=FALSE}
all_data$log_tenure <- log(all_data$tenure)
all_data$log_vantage <- log(all_data$Vantage_Avg)
all_data$log_income <- log(all_data$INCOME_INSIGHT_SCORE)


#Reference Categorys Changes#

# Ref Cat for Gender = Male
all_data$GENDER <- as.factor(all_data$GENDER)
all_data$GENDER <- relevel(all_data$GENDER, ref="Male")

# Ref Cat for Marriage = "Single"

all_data$MARRIAGE <- as.factor(all_data$MARRIAGE)
all_data$MARRIAGE <- relevel(all_data$MARRIAGE, ref="Single")


mod1 <- glm(Churn ~ tenure + kwh_solar + log_vantage, data = all_data, family = "binomial")
mod2 <- glm(Churn ~ tenure + kwh_solar + GENDER +  log_vantage + HOMEOWNER, data = all_data, family = "binomial")
mod3 <- glm(Churn ~ tenure + HOMEOWNER + log_income + EDUCATION, data = all_data, family = "binomial")
mod4 <- glm(Churn ~ tenure + kwh_solar + EDUCATION + log_income + MARRIAGE, data = all_data, family = "binomial")
mod5 <- glm(Churn ~ tenure  + GENDER + log_vantage  + OCCUPATION + HOMEOWNER, data = all_data, family = "binomial")
  
# summary(mod1)
# summary(mod2)
# summary(mod3)
# summary(mod4)
# summary(mod5)

```

#### Table 6: Churn Models

```{r echo=FALSE, message=FALSE, warning=FALSE, results='asis'}

stargazer(mod1,mod2,mod3, mod4, mod5, type = "html",  #we use html output to match our planned R Markdown output
     title = "Table 6: Churn Models")
```






From Table 6, it is noticeable how tenure length, here measured logarithmically transformed, is consistently measured as statistically significant. This effect is measured even when controlling for a combination of demographic and socioeconomic variables. Increased tenure lengths are associated with decreased probability of churn. For example, in Model 4, taking the odds ratio of the monthly tenure variable, a one month increase in tenure, controlling for solar kWh, Education, income and marital status, is associated with a 14% decrease in the odds of churning. This is supported by the logit curve in the Discussion section. 

This is one of the few statistically significant results observed in the analysis. The only other statistically significant variable was Gender. In model 5, Female account holders are associated 45% decrease in odds of churn, controlling for tenure, Vantage score, occupation and owning one's home. 

No other variables were associated with statistically significant effects on probability of churn.

```{r message=FALSE, warning=FALSE, include=FALSE}

# Odds Ratio Calculation used in above paragraph

## Tenure, Model 4

a = mod4$coefficients[2]

or_tenure_mod4 = exp(a[[1]])

change_one_month_tenure = (or_tenure_mod4 - 1)*100
change_one_month_tenure

## Gender, Model 5


a = mod5$coefficients[3]

or_tenure_mod5 = exp(a[[1]])

female_change = (or_tenure_mod5-1)
female_change

```




```{r echo=FALSE, message=FALSE, warning=FALSE}

mod1 <- glm(Default ~ tenure + kwh_solar + log_vantage, data = all_data, family = "binomial")
mod2 <- glm(Default ~ tenure + kwh_solar + GENDER +  log_vantage + HOMEOWNER, data = all_data, family = "binomial")
mod3 <- glm(Default ~ tenure + HOMEOWNER + log_income + EDUCATION, data = all_data, family = "binomial")
mod4 <- glm(Default ~ tenure + kwh_solar + EDUCATION + log_income + MARRIAGE, data = all_data, family = "binomial")
mod5 <- glm(Default ~ tenure  + GENDER + log_vantage  + OCCUPATION + HOMEOWNER, data = all_data, family = "binomial")

# summary(mod1)
# summary(mod2)
# summary(mod3)
# summary(mod4)
# summary(mod5)
```


#### Table 7: Default Models
```{r echo=FALSE, message=FALSE, warning=FALSE, results='asis'}

stargazer(mod1,mod2,mod3, mod4, mod5, type = "html",  #we use html output to match our planned R Markdown output
     title = "Table 7: Default Models")
```




# Secondary Analysis

The secondary analysis sought to determine if a measurable change can be determined in participant credit scores (Vantage Score). 

The graph below shows boxplots of each Vantage Score collected, at December 2020 and April 2022. 

```{r echo=FALSE, message=FALSE, warning=FALSE}
df <- final_joined %>%
  select(ParticipantID,`April 2022` = VANTAGE_V4_SCORE.x, `December 2020` = VANTAGE_V4_SCORE.y )

graph_temp <- reshape2::melt(df, id.var='ParticipantID') 


ggplot(graph_temp, aes(variable, value)) +
        geom_boxplot() +
        xlab("VANTAGE Score")
```

The histogram below plots the distribution of the scores by time period, and includes the standard scale and a logged comparison. 

```{r echo=FALSE, message=FALSE, warning=FALSE}




p1 <- ggplot(graph_temp, aes(value)) +
        geom_histogram(fill = "white", color = "grey30") +
        facet_wrap(~ variable) + ggtitle("Historgram of VANTAGE Score")

p2 <- ggplot(graph_temp, aes(value)) +
        geom_histogram(fill = "white", color = "grey30") +
        facet_wrap(~ variable) +
        scale_x_log10() + ggtitle("Historgram of VANTAGE Score, logged")

grid.arrange(p1, p2, nrow = 2)
```
Using a Welch t-test, we measure the difference in means between the two groups. From the below, the means of the two groups are not statistically signifcant; therefore we cannot claim that credit scores for individuals in the solar farms were observed to have changed. 

```{r echo=FALSE, message=FALSE, warning=FALSE}
t.test(value ~ variable, data = graph_temp)

```

```{r message=FALSE, warning=FALSE, include=FALSE}
# Same result as above, just avoiding the 'melt' function
t.test(final_joined$VANTAGE_V4_SCORE.x, final_joined$VANTAGE_V4_SCORE.y)

```




# Discussion


Primary:
- Higher Tenure -> lower likelihood
- Females less likely to churn than males
- no other observable effects, including rates by income, credit score, occupation, homeownership; meaning we didn't see any significnatly higher rates for populations who may have unfair biases and/or stereotypes in CS as being more likely to churn (LMI, Low-FICO, unscored/thin credit, etc)
- Default: not enough data, nothing more to really right here. 

Secondary:
- Credit Scores didn't change


The logit curves below show the relationship between the distribution of both tenure and days late with the probability of churning. The first graph shows that the longer the tenure, the lower the likelihood. The second graph shows that the days of late payments are generally not associated with likelihood of churn. 

```{r echo=FALSE, message=FALSE, warning=FALSE}

# Logit Curves - none significant except churn


ggplot(all_data, aes(x=tenure, y = Churn))+ geom_point(fill='red') + 
  stat_smooth(method="glm", method.args=list(family=binomial(link='logit')), se=FALSE)+theme_minimal()+ 
  scale_x_continuous(name="Months") 
  scale_y_continuous(name="Likelihood of Churn")
  
ggplot(all_data, aes(x=log_vantage, y = Churn))+ geom_point(fill='red') + 
  stat_smooth(method="glm", method.args=list(family=binomial(link='logit')), se=FALSE)+theme_minimal()+ 
  scale_x_continuous(name="Log VANTAGE Score") 
  scale_y_continuous(name="Likelihood of Churn")


ggplot(all_data, aes(x=log_income, y = Churn))+ geom_point(fill='red') + 
  stat_smooth(method="glm", method.args=list(family=binomial(link='logit')), se=FALSE)+theme_minimal()+ scale_x_continuous(name="Log Income Insight Score") +
  scale_y_continuous(name="Likelihood of Churn")

```



2. Secondary:
- No statistically significant effect was observed on credit score change after enrollment in solar farm. 


























