---
title: "SETO Data Analysis"
author: "Jacob Ford, Solstice Data Scientist"
date: "`r Sys.Date()`"
output: 
  rmdformats::readthedown:
    toc_depth: 5
    
knit: (function(inputFile, encoding) { 
        rmarkdown::render(inputFile, encoding = encoding, 
        output_file = file.path(dirname(inputFile), 'index.html')) })
---


<style>

# Input Files
{
   display: none;
 }
</style>



```{r message=FALSE, warning=FALSE, include=FALSE}

#Load packages we'll need throughout the entire document
library(readxl)
library(sf)
library(dplyr)
library(janitor)
library(dplyr)
library(tidycensus)
library(tidyverse)
options(tigris_use_cache = TRUE)
library(nngeo)
library(plotly)
library(kableExtra)
library(haven)
library(rms)
library(rmdformats)
library(ggplot2)
library(plotly)


options(scipen = 100)


```

```{r message=FALSE, warning=FALSE, include=FALSE}

## Change the below to match your local drive ##

input_folder <- '/Volumes/GoogleDrive/Shared drives/*Solstice | Low Income Inclusion/Projects/2019-2022 DOE SETO Contract Innovations/DOE SETO Project Implementation/Project Management/SETO Data Analysis/'



setwd(input_folder)


```



```{r message=FALSE, warning=FALSE, include=FALSE}

# 1. Community Solar Priorities

survey_name = paste(input_folder,"1. Community Solar Priorities/Qualtrics Survey/seto_survey/output/QUALTRICS_survey_data_all.xlsx", sep="")

survey <- read_excel(survey_name)

# 2. Developer & Financier

dev_input = paste(input_folder, "2. Developer and Financier/", sep="")

#dev_input ='/Volumes/GoogleDrive/Shared drives/*Solstice | Low Income Inclusion/Projects/2019-2022 DOE SETO Contract Innovations/DOE SETO Project Implementation/Project Management/Developer and Financier Survey/'




# 3. Churn & Default

cd_input = paste(input_folder, "3. Churn and Default/", sep="")


#cd_input = '/Volumes/GoogleDrive/Shared drives/*Solstice | Low Income Inclusion/Projects/2019-2022 DOE SETO Contract Innovations/DOE SETO Project Implementation/Project Management/Churn & Default Analysis'




```




# Introduction

This document contains a detailed description of the data analysis plan for Solstice Initiative’s research titled, “Product Innovation to Increase Low- to Moderate-Income Customers’ Adoption of Community Solar PV”.

This research will incorporate data from three distinct datasets: 

1. **Community Solar Priorities** research survey forms completed by respondents of varying demographics assessing the favorability of factorial test contracts and offering feedback on various aspects of the community solar product; 

2. **Developer and Financier Research** survey responses from developers and financiers to assess price sensitivities of language used in contract agreements; 

3. **Churn and Default Research** records of subscribers leaving programs and missing payments (“churn and default data”) provided by solar developers. 


















